Concurrent online sampling, for all, for free

### 2022-07-13 paper/main.tex
[wip] paper

    \subsection{ownership of Skips}
        % Recap
        In order to determine whether a tuple is a sample tuple or not a threads needs to own a Skip. To acquire a Skip the threads needs access the LoS. However, accessing the LoS requires synchronization and if the LoS contains only one Skip the thread needs to generate a new Skip.

@@

        % Actual acquiring
        The process of acquire a node from the LoS can be broken down into three steps. First the thread needs to read the head pointer and store the current value. Afterwards, the thread can use the index stored in the pointer to retrieve the node form the LoS. In the second step the thread needs to generate the successor for the Skip stored in the node and lastly the thread tries to writes the successor into the head of the LoS. If the update succeeds the thread now owns a new Skip and the head of the LoS is set to the successor Skip. On the other hand if the update fails, the thread needs to start over. However, in practice this is not a problem, because every time a thread fails to update the head another thread successfully updated the head and therefore overall progress is guaranteed. Figure \ref{fig:acquire} shows the three steps necessary to acquire a Skip and the retry loop.
        \begin{figure}[H]
            \includegraphics[height=2.4cm]{figure5.pdf}
            \caption{Flow diagram that shows the process of acquire a Skip from the LoS}
            \label{fig:acquire}
        \end{figure}

        % Generating the successor
        If the LoS only contains one node, the thread needs to generate a successor before takeing ownership of this node. In order to generate a Skip the thread needs to own a node. It could be the case that the thread already owns a node from a previous iteration or if the thread does not own a node it needs to retrieve one from the free list. Using a node and the current Skip at the head of the LoS the thread can generate the successor. However, it could also be the case that the LoS contains more then one node and no successor needs to be generated. But if the thread already owns a node, it needs to deallocate this node. Because a thread cannot own more then one node. Deallocateing a node refers to inserting the node into the free list. To summarize, there a two cases where a thread needs to allocate or deallocate a node: 
        \begin{itemize}
            \item Allocate: If the thread does not own a node and the LoS only contains one node
            \item Deallocate: If the thread owns a node and the LoS contains more the one node
        \end{itemize}

        % Returning a skip
        Hence an inserting thread could be terminated at any time it needs to be possible for a thread to return the Skip it is currently using. To return a skip

    \section{Example}
        % An example who the algorithm works (lar to one in the OG)
        This is an example.

>>>>

    \subsection{Ownership of Skips}
        % Recap
        In order to determine whether a tuple is a sample tuple or not a threads needs to own a Skip. To acquire a Skip the threads needs access the LoS. However, accessing the LoS requires synchronization and if the LoS contains only one Skip the thread needs to generate a new Skip.

@@

        % Actual acquiring
        The process of acquire a node from the LoS can be broken down into three steps. First the thread needs to read the head pointer and store the current value. Afterwards, the thread can use the index stored in the pointer to retrieve the node form the LoS. In the second step, the thread needs to generate the successor for the Skip stored in the node and lastly the thread tries to writes the successor into the head of the LoS. If the update succeeds the thread now owns a new Skip and the head of the LoS is set to the successor Skip. On the other hand, if the update fails, the thread needs to start over. However, in practice this is not a problem, because every time a thread fails to update the head, another thread successfully updated the head and therefore overall progress is guaranteed. Figure \ref{fig:acquire} shows the three steps necessary to acquire a Skip and the retry loop.
        \begin{figure}[H]
            \includegraphics[height=2.4cm]{figure5.pdf}
            \caption{Flow diagram that shows the process of acquire a Skip from the LoS}
            \label{fig:acquire}
        \end{figure}

        % Generating the successor
        If the LoS only contains one node, the thread needs to generate a successor before takeing ownership of this node. In order to generate a Skip the thread needs to own a node, because nodes can only be modified when they are owned by a thread. It could be the case that the thread already owns a node from a previous iteration. On the other hand, if the thread does not own a node it needs to retrieve one from the free list. Using a node and the current Skip at the head of the LoS the thread can generate the successor. However, it could also be the case that the LoS contains more then one node and no successor needs to be generated. But if the thread already owns a node, it needs to deallocate this node. Because a thread cannot own more then one node. Deallocateing a node refers to inserting the node into the free list. To summarize, there a two cases where a thread needs to allocate or deallocate a node: 
        \begin{itemize}
            \item Allocate: If the thread does not own a node and the LoS only contains one node
            \item Deallocate: If the thread owns a node and the LoS contains more the one node
        \end{itemize}

        % Returning a skip
        Since an inserting thread could be terminated at any time it needs to be possible for a thread to return the Skip it is currently using. To return a Skip, the Skip needs to be added to the LoS.

    \section{Example}
        % An example who the algorithm works (like the one in the OG)
        The following example will illustrate how the algorithm might be executed with two inserting threads $T0$ and $T1$. Figure \ref{fig:example} visualizes the state of all nodes and Skips used in this example. This example assumes that the sample as already been prefilled.
        \begin{figure}[H]
            \includegraphics[height=8cm]{figure6.pdf}
            \caption{The state of every Skip and node for the example}
            \label{fig:example}
        \end{figure}

        Initially all nodes are in the free list, except one node that contains the initial Skip (0). If $T0$ starts inserting it needs to acquire a Skip from the LoS. But because the LoS only contains one entry, $T0$ needs to generate a successor. However, $T0$ does not own a node and therefore it needs to retrieve one from the free list. Using this node and the initial Skip $T0$ can now generate the first Skip and write it to the head of the LoS. Afterwards $T0$ owns the initial Skip and the LoS contains the newly generated first Skip (1). When $T1$ also starts inserting it also needs to retrieve a Skip from the LoS. Since the LoS still contains just one Skip, $T1$ also needs to generate the second Skip and insert it into the LoS (2). In the meantime $T0$ reach a sample tuple and needs a new Skip to continue inserting. The LoS only contains one element, but this time $T0$ already owns a node from the previous iteration. Therefore, $T0$ can use its node and the Skip from the LoS to generate the third Skip. $T0$ can now swap the newly generated Skip with the second Skip, to take ownership of the second Skip (3). In this example $T1$ is terminated before it reaches a sample tuple. In consequence $T1$ needs to return its Skip to the LoS. The LoS now contains two element, Skip 1 that was previously owned by $T1$ and Skip 3 that was generated by $T0$ (4). At last $T0$ needs another Skip to continue inserting. This time $T0$ again owns a node from the previous iteration, but the LoS contains two elements. Therefore $T0$ does not need to generate a successor and can simply take ownership of the first Skip stored in the LoS. However, $T0$ cannot own two nodes and therefore $T0$ inserts the node of Skip into the free list (5).

        This example contains both allocation and deallocation. In step 1 $T0$ needs to allocate a node and in step 2 $T1$ needs to allocate a node. Because $T1$ returns its Skip the LoS contains two Skips after Step 4. Therefore $T0$ needs to deallocate the node it owns in Step 5.
